---
title: "Are 95% HPD intervals correct?"
output: html_notebook
---


# Inference without constraints
```{r}
d <- read.table("testHPD.csv", h=T)


```


```{r}
plot(d$true, d$low, pch=20, col=rgb(0,0,1,0.5), xlab="True dates", ylab="95HPD", ylim=c(0,60), xlim=c(0,60))
points(d$true, d$high, pch=20, col=rgb(1,0,0,0.5))
segments(d$true, d$low, d$true, d$high, col="black")
abline(a=0, b=1, col="grey", lwd=2)
```


# Inference with 5 constraints

```{r}
d <- read.table("testHPD_5Cons.csv", h=T)

```


```{r}
plot(d$true, d$low, pch=20, col=rgb(0,0,1,0.5), xlab="True dates", ylab="95HPD", ylim=c(0,60), xlim=c(0,60))
points(d$true, d$high, pch=20, col=rgb(1,0,0,0.5))
segments(d$true, d$low, d$true, d$high, col="black")
abline(a=0, b=1, col="grey", lwd=2)
```


# Are the parameters of the WNR model inferred well?
```{r}
d<- read.table("OutputDatingWNr/Cal_10_y_y_Cons_0_cons_BD_WNr_BL.log", h=T)
colnames(d)
```



# Are the parameters of the WNR model inferred well?

```{r}
summary(d$mean_rt)
```

```{r}
summary(d$var_rt)

```


# Other analysis, after fixing the rates


```{r}
d <- read.table("testHPD_FixedRates.csv", h=T)

```


```{r}
plot(d$true, d$low, pch=20, col=rgb(0,0,1,0.5), xlab="True dates", ylab="95HPD", ylim=c(0,60), xlim=c(0,60))
points(d$true, d$high, pch=20, col=rgb(1,0,0,0.5))
segments(d$true, d$low, d$true, d$high, col="black")
abline(a=0, b=1, col="grey", lwd=2)
```

I guess I did not fix the rates well... Let's check:


```{r}
d<- read.table("OutputDatingWNr/Cal_10_y_y_Cons_0_Fixed_rate_cons_BD_WNr_BL.log", h=T)
```


```{r}
summary(d$mean_rt)
```

```{r}
summary(d$var_rt)

```
Rates are completely wrong. I guess I don't really understand WN_r!

# Trying to run the analysis under the correct model of sequence evolution: HKY + Gamma20

```{r}
d <- read.table("testHPD_HKY.csv", h=T)

```


```{r}
plot(d$true, d$low, pch=20, col=rgb(0,0,1,0.5), xlab="True dates", ylab="95HPD", ylim=c(0,60), xlim=c(0,60))
points(d$true, d$high, pch=20, col=rgb(1,0,0,0.5))
segments(d$true, d$low, d$true, d$high, col="black")
abline(a=0, b=1, col="grey", lwd=2)
```


```{r}
d<- read.table("OutputDatingWNr/Cal_10_y_y_Cons_0_HKY_cons_BD_WNr_BL.log", h=T)
```

```{r}
summary(d$mean_rt)
```

```{r}
summary(d$var_rt)

```


# Trying simulation BD + JC and inference under the same model
```{r}
d <- read.table("SimuAndInfer/testHPD_SimuAndINfer.csv", h=T)

```



```{r}
plot(d$true, d$low, pch=20, col=rgb(0,0,1,0.5), xlab="True dates", ylab="95HPD", ylim=c(0,60), xlim=c(0,60))
points(d$true, d$high, pch=20, col=rgb(1,0,0,0.5))
segments(d$true, d$low, d$true, d$high, col="black")
abline(a=0, b=1, col="grey", lwd=2)
```


```{r}
plot(d$true, d$low, pch=20, col=rgb(0,0,1,0.5), xlab="True dates", ylab="95HPD", ylim=c(0,10), xlim=c(0,10))
points(d$true, d$high, pch=20, col=rgb(1,0,0,0.5))
segments(d$true, d$low, d$true, d$high, col="black")
abline(a=0, b=1, col="grey", lwd=2)
```



Basically, it seems like the BD parameters I chose are producing trees with nodes that are very close to the tips.


## Gergely Dos Reis analysis
```{r}
d <- read.table("TreesFromGergely/HPD.csv", h=T)

```



```{r}
plot(d$true, d$low, pch=20, col=rgb(0,0,1,0.5), xlab="True dates", ylab="95HPD", ylim=c(0,2), xlim=c(0,2))
points(d$true, d$high, pch=20, col=rgb(1,0,0,0.5))
segments(d$true, d$low, d$true, d$high, col="black")
abline(a=0, b=1, col="grey", lwd=2)
```



## Simu then infer, under UGAM model
```{r}
d <- read.table("SimuAndInferUGAM/testHPD.csv", h=T)

```



```{r}
plot(d$true, d$low, pch=20, col=rgb(0,0,1,0.5), xlab="True dates", ylab="95HPD", ylim=c(0,60), xlim=c(0,60))
points(d$true, d$high, pch=20, col=rgb(1,0,0,0.5))
segments(d$true, d$low, d$true, d$high, col="black")
abline(a=0, b=1, col="grey", lwd=2)
```


```{r}
plot(d$true, d$low, pch=20, col=rgb(0,0,1,0.5), xlab="True dates", ylab="95HPD", ylim=c(0,10), xlim=c(0,10))
points(d$true, d$high, pch=20, col=rgb(1,0,0,0.5))
segments(d$true, d$low, d$true, d$high, col="black")
abline(a=0, b=1, col="grey", lwd=2)
```




## Simu then infer, under UGAMr model
```{r}
d <- read.table("SimuAndInferUGAM/testHPD_UGAMr.csv", h=T)

```



```{r}
plot(d$true, d$low, pch=20, col=rgb(0,0,1,0.5), xlab="True dates", ylab="95HPD", ylim=c(0,60), xlim=c(0,60))
points(d$true, d$high, pch=20, col=rgb(1,0,0,0.5))
segments(d$true, d$low, d$true, d$high, col="black")
abline(a=0, b=1, col="grey", lwd=2)
```



## Simu then infer, under UGAMr model, with Bactrian moves
```{r}
d <- read.table("SimuAndInferUGAM/testHPD_Bactrian.csv", h=T)

```



```{r}
plot(d$true, d$low, pch=20, col=rgb(0,0,1,0.5), xlab="True dates", ylab="95HPD", ylim=c(0,60), xlim=c(0,60))
points(d$true, d$high, pch=20, col=rgb(1,0,0,0.5))
segments(d$true, d$low, d$true, d$high, col="black")
abline(a=0, b=1, col="grey", lwd=2)
```


## Simu then infer, under UGAMr model, with Bactrian moves and MC3

```{r}
d <- read.table("SimuAndInferUGAM/testHPD_MC3UGAMr.tsv", h=T)

```



```{r}
plot(d$true, d$low, pch=20, col=rgb(0,0,1,0.5), xlab="True dates", ylab="95HPD", ylim=c(0,60), xlim=c(0,60))
points(d$true, d$high, pch=20, col=rgb(1,0,0,0.5))
segments(d$true, d$low, d$true, d$high, col="black")
abline(a=0, b=1, col="grey", lwd=2)
```


```{r}
plot(d$true, d$low, pch=20, col=rgb(0,0,1,0.5), xlab="True dates", ylab="95HPD", ylim=c(0,10), xlim=c(0,10))
points(d$true, d$high, pch=20, col=rgb(1,0,0,0.5))
segments(d$true, d$low, d$true, d$high, col="black")
abline(a=0, b=1, col="grey", lwd=2)
```


##  Inference without constraints, under UGAMr model, with Bactrian moves

```{r}
d <- read.table("OutputDatingWNr/testHPD_UGAMr.tsv", h=T)

```



```{r}
plot(d$true, d$low, pch=20, col=rgb(0,0,1,0.5), xlab="True dates", ylab="95HPD", ylim=c(0,60), xlim=c(0,60))
points(d$true, d$high, pch=20, col=rgb(1,0,0,0.5))
segments(d$true, d$low, d$true, d$high, col="black")
abline(a=0, b=1, col="grey", lwd=2)
```


##  Inference without constraints, under UGAMr model, with Bactrian moves and MC3

```{r}
d <- read.table("OutputDatingWNr/testHPD_UGAMr_MC3.tsv", h=T)

```



```{r}
plot(d$true, d$low, pch=20, col=rgb(0,0,1,0.5), xlab="True dates", ylab="95HPD", ylim=c(0,60), xlim=c(0,60))
points(d$true, d$high, pch=20, col=rgb(1,0,0,0.5))
segments(d$true, d$low, d$true, d$high, col="black")
abline(a=0, b=1, col="grey", lwd=2)
```